---
title: "E2 - Experiment -Tuesday's Data"
author: "Jose Rico"
date: "9/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages needed
```{r message=FALSE}
library(readr)
library(tidyverse)
library(randomizr)
```

## Load Class Roster to Perform Randomization of Treatment
```{r}
roster <-read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSWYF1ho6YTKp7ntHl-9CV6HcgJYYqcT53MZUBBKGByhCxzJmUymJF4GWINHdAPMREUSj6IosD7E4Gc/pub?gid=0&single=true&output=csv")
```

## Create 2 Groups (Control and Treatment)

```{r}
N <- nrow(roster)
roster_ra <- roster %>% 
  mutate(group = complete_ra(N=N, conditions = c("eyes_closed","eyes_open")))
```

## Take a look at the data frame with `glimpse`
```{r}
glimpse(roster_ra)
```

##Check out the `summary` command of our data frame

```{r}
summary(roster_ra)
```
 
## Compare the new groups for equal proportions of our categorical variables

Example: look at distribution of gender within the groups
```{r}
#visual
ggplot(data = roster_ra, aes(x=group, fill=gender))+
  geom_bar()
#numerical (the 1 mean the % is calculated by the row)
  prop.table(table(roster_ra$group, roster_ra$gender),1)
```

## Practice visualization for additional categorical variables

Categorical Variable #2
```{r}
ggplot(data = roster_ra, aes(x=group, fill=factor(fitness_level)))+
  geom_bar()
```

Categorical Variable #3
```{r}

```

## Compare the groups with respect to the numerical variables

Example: Visualize height
```{r}
ggplot(roster_ra, aes(x=age, fill=group))+
  geom_dotplot()+
  facet_wrap(~group, nrow=2)
```
Example: Look at average height
```{r}
roster_ra %>% 
  group_by(group) %>% 
  summarise(mean=mean(height))
```

## Practice - analyze the other numerical variable

Visualize 
```{r}

```
Calculate averages 
```{r}

```

## Analysis of the experimental results

#Load in the data from the survey by publishing Google Sheet
```{r}
breath <-read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRXUa2Xo3hx6WnNo0WSu1OK_5oODLIxwooSFzteeVBZGXsRh_9UxqwY0nueUGRJLrza0wJSb6M5ZLcp/pub?gid=0&single=true&output=csv")
```

# Compare the Treatment to the Control
Looking at differences in breathing time between control and treatment groups, using dotplot of each group

```{r}
ggplot(data=breath, aes(x=time, fill=group))+
  geom_dotplot()+
  facet_wrap(~group, nrow=2)
```

Another way to see the difference is to use a boxplot summary graphic
```{r}
ggplot(data=breath, aes(x=group, y=time, color=group))+
  geom_boxplot()+
  geom_point(position="jitter")+
  coord_flip()
```

We can also compare averages time by looking at numerical summary broken down by group

```{r}
breath %>% 
  group_by(group) %>% 
  summarize(mean=mean(time))

```

**Group Discusssion prompt: What do you notice?**

# Looking deeper into the categorical variables for differences within factors

Do we see any difference across categorical variable from control to treatment? 

Example: differences within **gender** across the treatment/control

Numerical Summaries (median and mean)
```{r}
breath %>% 
  group_by(gender, group) %>% 
  summarize(mean=mean(time))
```

Visualize the results with a dotplot

```{r}
ggplot(data=breath, aes( x=time, fill=practiced))+
  geom_dotplot()+
  facet_wrap(group~practiced, nrow=2)
```

How is this different?  When would I want this instead?
```{r}
ggplot(data=breath, aes( x=time, fill=group))+
  geom_dotplot()+
  facet_wrap(gender~group, nrow=2)
```

## Additional Practice
Now let's try to modify the above analysis for other categorical variables.
